---
title: "Random Forest Modeling and Prediction with Two models; Frequency and Severity"
author: "greg"
date: "November 8, 2016"
output: html_document
---
## Plan

```{r}
#Step 0 importing data and libraries
#Step 1.1 Facet Engineering and Heuristic Factors for train (Note: For next step, Copy into new file, ctrl f, repalce all "train" with test)
#Step 1 Facet Engineering and Heuristic Factors for test
#Step 2 Factor (train, test)
#Step 3 SVM 
#Step 4 GBM SOLO
#Step 5 (part 1) Logistic Regression to predict probability of claim
#Step 5 (part 2) GMB to predict cost
#Step 5 (part 3) Expected Value
```
      
## Step 0 Importing Data
```{r}
# importing training data
train.df<-read.csv("Kangaroo_train.csv")

# importing the validation data for test data
test.df<-read.csv("Kangaroo_valid.csv")

# importing validation data with actual values
repsonse<-read.csv("Kangaroo_valid_w_response.csv")

# check for na in data
any(is.na(train.df))

# import libraries
library(dplyr)
library(ggplot2)
library('reshape2')
library('gbm')
library('ROCR')

library("e1071")
library("class")
library("ranger")

library("caret")
library("glmnet")
library("ROCR")

#Commented out this handy sample. Use this code to remove na if I encounter them during facet engineering
#na<-is.na(train.df$new)
#train.df$new[na]<-0
```
## Facet Engineering Plan

-veh_value categorized

!!!!!Exposure!!!!!
looking at a (claim cost bin/exposure bin)we can get a better idea of how much the respective bin cost
looking at a (claim count bin/exposure bin)we can get a better idea of how likely a claim is in respective bin

-cost/exposure_relativity = (cost total of bin facet/exposure total of bin facet)/(average pure premium)
                                           = (cost total of bin facet/exposure total of bin facet)/pp
                                           
    Lets represent this as pr for pure premium relativity                             
                                           
    Note: Bins will be created through cross variable analysis
    gender&dr_age
    ie.
    -cost_male_age1_relativity
    -cost_female_age1_relativity
    -cost_female_age2_relativity
   
   veh_age&veh_value
    
These factors show us how much better or worse a group is than the average. It will be used to differentiate. I will multiply the final prediction by these values. This will boost the gini coeficient. 

-count/exposure_relativity = (claim total of bin facet/exposure total of bin facet)/(total dataset claim_count/total exposure)

      Lets represent this as cr for count relativity 
      
## Step 1 FACET ENGINEERING FOR TRAIN
```{r}
############################################################################
#             NEW VEH_BOD CATEGORIES
############################################################################

#new variable for sport
s<-train.df$veh_body=="COUPE"|train.df$veh_body=="CONVT"|train.df$veh_body=="HDTOP"|train.df$veh_body=="RDSTR"
#new variable for bus
b<-train.df$veh_body=="BUS"|train.df$veh_body=="MCARA"|train.df$veh_body=="MIBUS"|train.df$veh_body=="PANVN"

#change veh_body from factor to character
train.df$veh_body<-as.character(train.df$veh_body)

#add new factor(grouped categories) as string called "SPORT"
train.df$veh_body[s]<-"SPORT"
#add new factor(grouped categories) as string called "BUS"
train.df$veh_body[b]<-"BUS"

############################################################################
#             VEH_VAL CHANGE TO CATEGORIES
############################################################################

#lets multiply by 100 to make the ranges easier to work with
train.df$veh_value<-train.df$veh_value*100
#now that we have integers we need to convert the data
train.df$veh_value<-as.integer(train.df$veh_value)
#ifelse&mutate to change the variable from numeric to categorical
train.df<-mutate(train.df, veh_value = ifelse(veh_value %in% 00:99, "A",
                                     ifelse(veh_value %in% 100:199, "B",
                                            ifelse(veh_value %in% 200:299, "C",
                                                   ifelse(veh_value %in% 300:399, "D",
                                                          ifelse(veh_value %in% 400:499, "E", "F"))))))

############################################################################
#             NEW PR AND CR VARIABLES
############################################################################

#pure premium average
j1<-sum(train.df$claim_cost)
j2<-sum(train.df$exposure)
pp<-j1/j2

#pr for males/dr_age
M_1_pr<-(sum(train.df$claim_cost[train.df$gender=="M"&train.df$dr_age==1])/sum(train.df$exposure[train.df$gender=="M"&train.df$dr_age==1]))/pp
M_2_pr<-(sum(train.df$claim_cost[train.df$gender=="M"&train.df$dr_age==2])/sum(train.df$exposure[train.df$gender=="M"&train.df$dr_age==2]))/pp
M_3_pr<-(sum(train.df$claim_cost[train.df$gender=="M"&train.df$dr_age==3])/sum(train.df$exposure[train.df$gender=="M"&train.df$dr_age==3]))/pp
M_4_pr<-(sum(train.df$claim_cost[train.df$gender=="M"&train.df$dr_age==4])/sum(train.df$exposure[train.df$gender=="M"&train.df$dr_age==4]))/pp
M_5_pr<-(sum(train.df$claim_cost[train.df$gender=="M"&train.df$dr_age==5])/sum(train.df$exposure[train.df$gender=="M"&train.df$dr_age==5]))/pp
M_6_pr<-(sum(train.df$claim_cost[train.df$gender=="M"&train.df$dr_age==6])/sum(train.df$exposure[train.df$gender=="M"&train.df$dr_age==6]))/pp


#pr for females/dr_age
F_1_pr<-(sum(train.df$claim_cost[train.df$gender=="F"&train.df$dr_age==1])/sum(train.df$exposure[train.df$gender=="F"&train.df$dr_age==1]))/pp
F_2_pr<-(sum(train.df$claim_cost[train.df$gender=="F"&train.df$dr_age==2])/sum(train.df$exposure[train.df$gender=="F"&train.df$dr_age==2]))/pp
F_3_pr<-(sum(train.df$claim_cost[train.df$gender=="F"&train.df$dr_age==3])/sum(train.df$exposure[train.df$gender=="F"&train.df$dr_age==3]))/pp
F_4_pr<-(sum(train.df$claim_cost[train.df$gender=="F"&train.df$dr_age==4])/sum(train.df$exposure[train.df$gender=="F"&train.df$dr_age==4]))/pp
F_5_pr<-(sum(train.df$claim_cost[train.df$gender=="F"&train.df$dr_age==5])/sum(train.df$exposure[train.df$gender=="F"&train.df$dr_age==5]))/pp
F_6_pr<-(sum(train.df$claim_cost[train.df$gender=="F"&train.df$dr_age==6])/sum(train.df$exposure[train.df$gender=="F"&train.df$dr_age==6]))/pp

#Now lets turn pr into a variable for our data set
train.df$pr<-0
train.df$pr[train.df$gender=="M"&train.df$dr_age==1]<-M_1_pr
train.df$pr[train.df$gender=="M"&train.df$dr_age==2]<-M_2_pr
train.df$pr[train.df$gender=="M"&train.df$dr_age==3]<-M_3_pr
train.df$pr[train.df$gender=="M"&train.df$dr_age==4]<-M_4_pr
train.df$pr[train.df$gender=="M"&train.df$dr_age==5]<-M_5_pr
train.df$pr[train.df$gender=="M"&train.df$dr_age==6]<-M_6_pr

train.df$pr[train.df$gender=="F"&train.df$dr_age==1]<-F_1_pr
train.df$pr[train.df$gender=="F"&train.df$dr_age==2]<-F_2_pr
train.df$pr[train.df$gender=="F"&train.df$dr_age==3]<-F_3_pr
train.df$pr[train.df$gender=="F"&train.df$dr_age==4]<-F_4_pr
train.df$pr[train.df$gender=="F"&train.df$dr_age==5]<-F_5_pr
train.df$pr[train.df$gender=="F"&train.df$dr_age==6]<-F_6_pr

#cr average
j3<-sum(train.df$claim_count)
j4<-sum(train.df$exposure)
cr<-j3/j4

#cr for males/dr_age
M_1_cr<-(sum(train.df$claim_count[train.df$gender=="M"&train.df$dr_age==1])/sum(train.df$exposure[train.df$gender=="M"&train.df$dr_age==1]))/cr
M_2_cr<-(sum(train.df$claim_count[train.df$gender=="M"&train.df$dr_age==2])/sum(train.df$exposure[train.df$gender=="M"&train.df$dr_age==2]))/cr
M_3_cr<-(sum(train.df$claim_count[train.df$gender=="M"&train.df$dr_age==3])/sum(train.df$exposure[train.df$gender=="M"&train.df$dr_age==3]))/cr
M_4_cr<-(sum(train.df$claim_count[train.df$gender=="M"&train.df$dr_age==4])/sum(train.df$exposure[train.df$gender=="M"&train.df$dr_age==4]))/cr
M_5_cr<-(sum(train.df$claim_count[train.df$gender=="M"&train.df$dr_age==5])/sum(train.df$exposure[train.df$gender=="M"&train.df$dr_age==5]))/cr
M_6_cr<-(sum(train.df$claim_count[train.df$gender=="M"&train.df$dr_age==6])/sum(train.df$exposure[train.df$gender=="M"&train.df$dr_age==6]))/cr


#cr for females/dr_age
F_1_cr<-(sum(train.df$claim_count[train.df$gender=="F"&train.df$dr_age==1])/sum(train.df$exposure[train.df$gender=="F"&train.df$dr_age==1]))/cr
F_2_cr<-(sum(train.df$claim_count[train.df$gender=="F"&train.df$dr_age==2])/sum(train.df$exposure[train.df$gender=="F"&train.df$dr_age==2]))/cr
F_3_cr<-(sum(train.df$claim_count[train.df$gender=="F"&train.df$dr_age==3])/sum(train.df$exposure[train.df$gender=="F"&train.df$dr_age==3]))/cr
F_4_cr<-(sum(train.df$claim_count[train.df$gender=="F"&train.df$dr_age==4])/sum(train.df$exposure[train.df$gender=="F"&train.df$dr_age==4]))/cr
F_5_cr<-(sum(train.df$claim_count[train.df$gender=="F"&train.df$dr_age==5])/sum(train.df$exposure[train.df$gender=="F"&train.df$dr_age==5]))/cr
F_6_cr<-(sum(train.df$claim_count[train.df$gender=="F"&train.df$dr_age==6])/sum(train.df$exposure[train.df$gender=="F"&train.df$dr_age==6]))/cr


#Finally make cr into a variable for our data
train.df$cr[train.df$gender=="M"&train.df$dr_age==1]<-M_1_cr
train.df$cr[train.df$gender=="M"&train.df$dr_age==2]<-M_2_cr
train.df$cr[train.df$gender=="M"&train.df$dr_age==3]<-M_3_cr
train.df$cr[train.df$gender=="M"&train.df$dr_age==4]<-M_4_cr
train.df$cr[train.df$gender=="M"&train.df$dr_age==5]<-M_5_cr
train.df$cr[train.df$gender=="M"&train.df$dr_age==6]<-M_6_cr

train.df$cr[train.df$gender=="F"&train.df$dr_age==1]<-F_1_cr
train.df$cr[train.df$gender=="F"&train.df$dr_age==2]<-F_2_cr
train.df$cr[train.df$gender=="F"&train.df$dr_age==3]<-F_3_cr
train.df$cr[train.df$gender=="F"&train.df$dr_age==4]<-F_4_cr
train.df$cr[train.df$gender=="F"&train.df$dr_age==5]<-F_5_cr
train.df$cr[train.df$gender=="F"&train.df$dr_age==6]<-F_6_cr


```

## Step 1.1 FACET ENGINEERING FOR TEST
```{r}
############################################################################
#             NEW VEH_BOD CATEGORIES 
############################################################################

#new variable for sport
s<-test.df$veh_body=="COUPE"|test.df$veh_body=="CONVT"|test.df$veh_body=="HDTOP"|test.df$veh_body=="RDSTR"
#new variable for bus
b<-test.df$veh_body=="BUS"|test.df$veh_body=="MCARA"|test.df$veh_body=="MIBUS"|test.df$veh_body=="PANVN"

#change veh_body from factor to character
test.df$veh_body<-as.character(test.df$veh_body)

#add new factor(grouped categories) as string called "SPORT"
test.df$veh_body[s]<-"SPORT"
#add new factor(grouped categories) as string called "BUS"
test.df$veh_body[b]<-"BUS"

############################################################################
#             VEH_VAL CHANGE TO CATEGORIES
############################################################################

#lets multiply by 100 to make the ranges easier to work with
test.df$veh_value<-test.df$veh_value*100
#now that we have integers we need to convert the data
test.df$veh_value<-as.integer(test.df$veh_value)
#ifelse&mutate to change the variable from numeric to categorical
test.df<-mutate(test.df, veh_value = ifelse(veh_value %in% 00:99, "A",
                                     ifelse(veh_value %in% 100:199, "B",
                                            ifelse(veh_value %in% 200:299, "C",
                                                   ifelse(veh_value %in% 300:399, "D",
                                                          ifelse(veh_value %in% 400:499, "E", "F"))))))


############################################################################################
#             NEW PR AND CR VARIABLES 
############################################################################################

#I use the relativity values engineered from the bins in the training data to construct new variables for the test set
#Now lets turn pr into a variable for our data set
test.df$pr<-0
test.df$pr[test.df$gender=="M"&test.df$dr_age==1]<-M_1_pr
test.df$pr[test.df$gender=="M"&test.df$dr_age==2]<-M_2_pr
test.df$pr[test.df$gender=="M"&test.df$dr_age==3]<-M_3_pr
test.df$pr[test.df$gender=="M"&test.df$dr_age==4]<-M_4_pr
test.df$pr[test.df$gender=="M"&test.df$dr_age==5]<-M_5_pr
test.df$pr[test.df$gender=="M"&test.df$dr_age==6]<-M_6_pr

test.df$pr[test.df$gender=="F"&test.df$dr_age==1]<-F_1_pr
test.df$pr[test.df$gender=="F"&test.df$dr_age==2]<-F_2_pr
test.df$pr[test.df$gender=="F"&test.df$dr_age==3]<-F_3_pr
test.df$pr[test.df$gender=="F"&test.df$dr_age==4]<-F_4_pr
test.df$pr[test.df$gender=="F"&test.df$dr_age==5]<-F_5_pr
test.df$pr[test.df$gender=="F"&test.df$dr_age==6]<-F_6_pr

#Finally make cr into a variable for our data
test.df$cr[test.df$gender=="M"&test.df$dr_age==1]<-M_1_cr
test.df$cr[test.df$gender=="M"&test.df$dr_age==2]<-M_2_cr
test.df$cr[test.df$gender=="M"&test.df$dr_age==3]<-M_3_cr
test.df$cr[test.df$gender=="M"&test.df$dr_age==4]<-M_4_cr
test.df$cr[test.df$gender=="M"&test.df$dr_age==5]<-M_5_cr
test.df$cr[test.df$gender=="M"&test.df$dr_age==6]<-M_6_cr

test.df$cr[test.df$gender=="F"&test.df$dr_age==1]<-F_1_cr
test.df$cr[test.df$gender=="F"&test.df$dr_age==2]<-F_2_cr
test.df$cr[test.df$gender=="F"&test.df$dr_age==3]<-F_3_cr
test.df$cr[test.df$gender=="F"&test.df$dr_age==4]<-F_4_cr
test.df$cr[test.df$gender=="F"&test.df$dr_age==5]<-F_5_cr
test.df$cr[test.df$gender=="F"&test.df$dr_age==6]<-F_6_cr
```

## Step 2 FACTORING
```{r}
#factoring train
train.df$dr_age<-factor(train.df$dr_age)
train.df$veh_age<-factor(train.df$veh_age)
train.df$veh_body<-factor(train.df$veh_body)
train.df$veh_value<-factor(train.df$veh_value)
#factoring test
test.df$dr_age<-factor(test.df$dr_age)
test.df$veh_age<-factor(test.df$veh_age)
test.df$veh_body<-factor(test.df$veh_body)
test.df$veh_value<-factor(test.df$veh_value)
```

## Step 3 SVM MODEL
```{r, eval=F}
# making a new train data set to work with for modeling
# Removing some response variables we no longer need

factor(train.df$claim_count)
train<-train.df[,-c(1,9,11)]

#making the new test set for prediction
test<-test.df[,-c(1)]

model_svm<-svm(claim_count~.,data=train,kernel="linear",cost=.1,scale=FALSE)
print(model_svm)
```

## Step 4 GBM SOLO MODEL
```{r, eval=F}
# making a new train data set to work with for modeling
# Removing some response variables we no longer need
train<-train.df[,-c(1,9,10)]

#making the new test set for prediction
test<-test.df[,-c(1)]

#Model/Prediction Freq Using Random Forest method="Ranger"
# Fit random forest: model
model_gbm1 <- train(
  claim_cost~.,
  tuneLength = 5,
  data = train, method = "gbm",
  trControl = trainControl(method = "cv", number = 5, verboseIter = TRUE)
)
print(model_gbm1)
summary(model_gbm1)
plot(model_gbm1)


#predict
premium_prediction<-predict(model_gbm1, test, type="raw")
#making data frame for predictions
pure_premium<-data.frame(id=test.df$id,premium_prediction)
#write the data.frame as csv
write.csv(pure_premium, file = "GBM1_predictions.csv")
```

## Step 5 (part1) GLM for claim probability
```{r}
# Removing some response variables we no longer need
train<-train.df[,-c(1,10,11)]

#making the new test set for prediction
test<-test.df[,-c(1)]

#the logistic model
model_glm<-glm(formula = claim_ind ~ ., family = binomial(link = "logit"), 
    data = train)

#predict probability of claim
claim_prediction<-predict(model_glm, test, type="response")
```

## Step 5 (part2) GBM for claim_cost, without claim_count=0
```{r}

#filter zero claim with dplyr

train0 <- dplyr::filter(train.df, claim_count != 0)

# Removing some response variables we no longer need

train<-train0[,-c(1,9,10)]

#making the new test set for prediction

test<-test.df[,-c(1)]

model_gbm2 <- train(
  claim_cost~.,
  tuneLength = 5,
  data = train, method = "gbm",
  trControl = trainControl(method = "cv", number = 5, verboseIter = TRUE)
)

#predicting cost

premium_prediction<-predict(model_gbm2, test, type="raw")

```

## Step 5 (part3) Multiply resulting
```{r}
#multiply prob by the predicted cost for an expected value

pure_premium_combo<-data.frame(test.df[,1],premium_prediction*claim_prediction, repsonse[,11])

#write.csv(pure_premium, file = "Hold_predictions1_JEBG.csv")
```